---
title: "Quick start to model fitting with brms"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(brms)
library(learnB4SS)
data("dataset")
brm_msg <- read_lines(system.file("extdata/brm-sampling.txt", package = "learnB4SS"))
```


## Fit your first Bayesian regression

### Taking the leap

<!-- TODO: a few sentences about what skills you need and what you already know (like lme4 syntax, etc...) -->

Welcome! Here...

<!-- <a href="#" data-toggle="tooltip" title="Hooray!">Hover over me</a> -->
<!-- <a href="#" data-toggle="popover" data-trigger="hover" title="Popover Header" data-content="Some content inside the popover">Toggle popover</a> -->

### Dance to the tune 

Since the complexity of Bayesian statistics can be intimidating at first, we would like to show you, in fact, how straightforward it is to fit a Bayesian regression with brms.

To this aim, we will use a dataset from .

```{r dataset}
dataset
```

The `dataset` dataframe contains .

You can find more info with `?dataset`.

What we want to know is whether .
In the frequentist world, you would proceed by fitting a linear regression, like:

```{r attitude-lm, exercise=TRUE}
attitude_lm <- lm(f0mn ~ attitude, data = dataset)
```

If you wanna turn Bayesian, it's easy!
The next section will show you how.

### Fit the model with `brm()`

If you want to run a Bayesian regression model, just replace `lm()` with `brm()` from the brms package!

```{r include=FALSE}
brm <- function(formula, data) {
  for (line in brm_msg) {
    cat(line)
    cat("\n")
  }
}
```


```{r attitude-bm, exercise=TRUE}
attitude_bm <- brm(f0mn ~ attitude, data = dataset)

```



## Exercises examples

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```


<!-- <script> -->
<!-- $(document).ready(function(){ -->
<!--   $('[data-toggle="popover"]').popover();  -->
<!-- }); -->
<!-- $(document).ready(function(){ -->
<!--   $('[data-toggle="tooltip"]').tooltip();  -->
<!-- }); -->
<!-- </script> -->
