---
title: "Application to regression II - The prior and Bayesian updating"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Application to regression II - The prior and Bayesian updating}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Load data

Let's load the data we are going to use.

```{r setup}
library(brms)
library(learnB4SS)
data("polite")
```


# Which priors?

Depending on the type of model you are running and the nature of the included predictors, you will have to specify a certain set of priors.

You can easily learn which priors you can specify in a model by using the function `brms::get_prior()`. You will try it out in a second.

We are trying to find out what kind of effect politeness has (if any) on articulation rate.

As a first approximation, we could start with the following model:

```r 
# Don't run me yet!
m1 <- brm(
  formula = articulation_rate ~ attitude,
  data = polite
)
```

To find out which priors are available in this model, you just need to specify the model formula *and* the data in the `get_prior()` function.

Go ahead and fill in the following code snippet and run it when you're done to get a list of priors!
You'll learn how to interpret the output in the next section.

```{r get-prior}
get_prior(
 formula = ...,
 data = ...
)
```

# Understand the syntax

When you run `get_prior()`, a table is returned with a few columns.
Each row corresponds to one prior (and to one model parameter).

For now, you can focus just on the first three (`prior`, `class`, `coef`).

## `prior`

The column `prior` tells you which prior probability distributions are set as default by brms.
For our model, the first two default priors are `(flat)`, i.e. uniform distributions (all values are equally probable).
The other two priors are Student-*t* distributions.
(more on prior specification below).

But we will want to change them all to something more appropriate (we will talk about what makes an appropriate prior tomorrow).

## `class`

The columns `class` and `coef` are the ones you should check.
`class` indicates the type of model parameter: `b` indicates a slope term, `Intercept` the model intercept, and `sigma` the model residual standard deviation.

## `coef`

The third column, `coef` indicates to which model predictor the prior is assigned.
In our model above, there's only one predictor (`attitude`), and that's what you can see in the `get_prior()` output.

Since `attitude` has only two levels (`inf` and `pol`), we only have one parameter for the second level `pol` (the level `inf` is included in the `Intercept`).

# Specify your priors

It's quite easy to specify priors with brms.
You just need to use the function `prior()`!

Since we need to specify more than one prior, we need multiple calls to `prior()`, concatenated with `c()`.
Like in the following example:

```r
my_priors <- c(
  prior(...),
  prior(...),
  prior(...)
)
```

Within the function `prior()`, you need to specify the prior distribution with its parameters, the prior class (`class`), and the prior coefficient (`coef`) if any.

For our simple model above, we need three priors:

- The prior of the `Intercept`.
- The prior of the slope for `attitudepol`.
- The prior for the standard deviation `sigma`.

Remember that `get_prior()` also lists class-generic priors (in our case, the prior for `class = b`), but you don't need to specify those if you specify the prior for the coefficient (in our case, `class = b, coef = attitudepol`).

We'll give you a head start by telling you which priors we can use in this model:

- `Intercept`: a Normal distribution with mean = 0 and sd = 15.
- `attitudepol`: a Normal distribution with mean = 0 and sd = 10.
- `sigma`: a Half-Cauchy distribution with mean = 0 and sd = 1.

Go ahead and fill in the following code snippet with the right parameters and arguments:

```{r set-priors}
my_priors <- c(
  prior(normal(..., ...), class = Intercept),
  prior(normal(..., ...), class = b, coef = ...),
  prior(cauchy(0, ...), class = ...)
)
```

And now let's fit the model using our priors!

```{r fit}
m1 <- brm(
  articulation_rate ~ attitude,
  data = polite,
  prior = my_priors
)
```

