---
title: "How to install the learnB4SS package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to install the learnB4SS package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

All the workshop materials can be conveniently installed as an R package.

To install the learnB4SS package, you need to install the remotes package first, if not already installed.

```{r remotes}
install.packages("remotes")
```

Then, simply ran the following line.

```{r learnb4ss}
remotes::install_github("learnB4SS/learnB4SS@v1.0.0")
```

🎉 You are now ready for your Bayesian adventure!

## Usage

During the workshop, you will be able to follow along by attaching the package learnB4SS:

```{r attach}
library(learnB4SS)
```

Now you can use the function `open_slides()` to open the session slides in your browser and `open_exercise()` to open the exercise files.

```{r open}
# Open slides of session 00
open_slides(00)

# Open exercise of session 03
open_exercise(03)
```

We will use the `polite` data throughout the workshop.

You can load the data with:

```{r data}
data("polite")

# Learn about the data
?polite
```

