---
title: "Understand and explore priors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Understand and explore priors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
library(learnB4SS)
data(emotion)
```

# Overview

This vignette digs a bit deeper into priors for different types of outcome/response variables.
We will cover outcome variables that follow these distributions: normal/Gaussian, log-normal, binomial/Bernoulli, poisson, and beta.


# Normal/Gaussian

Let's start off with an outcome variable distributed according to a normal/Gaussian distribution.

In fact, we hardly ever work with truly normally distributed variables.
So for this example we will look at a data set of lexical emotional valence.

```{r}
emotion
```

The `emotion` tibble contains a list of 1000 English words and their emotional valence.
Emotional valence is given as a number between `-4` and `+4`, 
